---
title: "EDA"
output: github_document
date: "2024-10-03"
---

```{r setup, include=FALSE}
library(tidyverse)

```

adding in month=floor_date() function
```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USW00022534", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2021-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = case_match(
      id, 
      "USW00094728" ~ "CentralPark_NY", 
      "USW00022534" ~ "Molokai_HI",
      "USS0023B17S" ~ "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10,
    month= floor_date(date, unit= "month")) |>
  select(name, id, everything())
```

Let's make some plots 

```{r}
weather_df |> 
  ggplot(aes(x=prcp)) +
  geom_histogram()
```

showing real datapoints that had abnormally large rainfalls. do we care about outliers?
```{r}
weather_df |> 
  filter(prcp>1000)
```

```{r}
weather_df |> 
  filter(tmax>20, tmax<30) |> 
  ggplot(aes(x=tmin, y=tmax, color=name, shape=name)) +
  geom_point()
```

## group_by()

tells you how many groups of whatever variable you're putting in
```{r}
weather_df |> 
  group_by(name)
```

counting stuff

tells us how many obs we have and how many distinct months we have
```{r}
weather_df |> 
  group_by(name) |> 
  summarize(
      n_obs=n(),
      n_dist=n_distinct(month))
```

grouping by name and month
```{r}
weather_df |> 
  group_by(name, month) |> 
  summarize(
      n_obs=n())
```

counting how many obs with count()
```{r}
weather_df |> 
  count(name)
```


## 2x2 

```{r}
weather_df |> 
  drop_na(tmax) |> 
  filter(name !="Molokai_HI") |> 
  mutate(
    cold = case_when(
      tmax <5 ~ "cold",
      tmax >=5 ~ "not_cold"
    )
  ) |> 
  group_by(name, cold) |> 
  summarize(count = n())
```
could pivot wider above to make a 2x2



janitor package tabyl() also produces a 2x2 (DO NOT USE table())
```{r}
weather_df |> 
  drop_na(tmax) |> 
  filter(name !="Molokai_HI") |> 
  mutate(
    cold = case_when(
      tmax <5 ~ "cold",
      tmax >=5 ~ "not_cold"
    )
  ) |> 
  janitor::tabyl(name, cold)
```

## general numeric summaries

```{r}

```



