---
title: "Vis II"
output: github_document
date: "2024-10-01"
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggridges)
library(patchwork)
```

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USW00022534", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2021-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = case_match(
      id, 
      "USW00094728" ~ "CentralPark_NY", 
      "USW00022534" ~ "Molokai_HI",
      "USS0023B17S" ~ "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())
```

Making out first plot :)

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) +
geom_point()
```
point says put points for all variables--making a scatterplot for us. aes is our aesthetics

```{r}
weather_df |> 
  ggplot(aes (x =tmin, y=tmax)) + 
  geom_point()
```
can also do it this way

```{r}
ggp_weather_scatterplot=
  weather_df |> 
   ggplot(aes (x =tmin, y=tmax)) + 
  geom_point()

ggp_weather_scatterplot
```

check why some rows are missing...
```{r}
weather_df |> 
  filter(is.na(tmax))
```
telling us there are 17 rows missing tmax


## Fancier scatterplots!

```{r}
weather_df |> 
  ggplot(aes(x=tmin, y=tmax, color=name)) +
  geom_point(alpha=.3, size=.8) + 
  geom_smooth(se = FALSE)
```
alpha makes points more transparent

Where you define aesthetics can matter--now putting color=name below and wont be applied to ggplot

```{r}
weather_df |> 
  ggplot(aes(x=tmin, y=tmax)) +
  geom_point(aes(color=name), alpha=.3, size=.8) + 
  geom_smooth(se = FALSE)
```

use faceting real quick

```{r}
weather_df |> 
  ggplot(aes(x=tmin, y=tmax, color=name)) +
  geom_point(alpha=.3) +
  geom_smooth(se= FALSE) + 
  facet_grid(. ~ name)
```
for faced grid . (nothing) goes in rows and name in columns. can reverse it too

Lets make a somewhat more interesting scatterplot 

```{r}
weather_df |> 
  ggplot(aes(x=date, y=tmax, color=name, size = prcp)) +
  geom_point(alpha=.3) +
  geom_smooth(se= FALSE) +
  facet_grid(. ~ name)
```

Learning assessment--need to manipulate data a little bit first

```{r}
weather_df |> 
  filter(name == "CentralPark_NY") |> 
  mutate(
    tmax_fahr = tmax + (9/5) + 32,
    tmin_fahr = tmin + (9/5) + 32) |> 
  ggplot(aes(x=tmin_fahr, y=tmax_fahr)) +
  geom_point() +
  geom_smooth(method= "lm", se=FALSE)
```

## Small Things

```{r}
weather_df |> 
  ggplot(aes(x=date, y=tmax,)) +
  geom_point(aes(color=name), alpha=.3, size=.8) +
  geom_smooth(se= FALSE) 
```

```{r}
weather_df |> 
  ggplot(aes(x=tmin, y=tmax)) +
  geom_hex()
```


```{r}
weather_df |> 
  ggplot(aes(x=tmin, y=tmax)) +
  geom_point(color= "blue")
```

## Univeriate plots

```{r}
weather_df |> 
  ggplot(aes(x=tmin)) + 
  geom_histogram()
```

```{r}
weather_df |> 
  ggplot(aes(x=tmin, fill=name)) + 
  geom_histogram(position = "dodge")
```
position command puts bars next to each other rather than on top of eachother

how would i fix this? 

maybe facet?
```{r}
weather_df |> 
  ggplot(aes(x=tmin, fill=name)) +
  geom_histogram() +
  facet_grid(. ~ name)
```

maybe a density plot?
```{r}
weather_df |> 
  ggplot(aes(x=tmin, fill=name)) +
  geom_density(alpha=.3)
```

boxplot
```{r}
weather_df |> 
  ggplot(aes(x=name, y=tmin, fill= name)) +
  geom_boxplot()
```
fill will give color to name

violin plots

```{r}
weather_df |> 
  ggplot(aes(x=name, y=tmin, fill= name)) +
  geom_violin()
```

ridge plot

```{r}
weather_df |> 
  ggplot(aes(x=tmin, y=name)) +
  geom_density_ridges()
```

## Learning Assessment

```{r}
weather_df |> 
  ggplot(aes(x=prcp, y=name, fill=name)) +
 geom_boxplot()
```

```{r}
weather_df |> 
  ggplot(aes(x=prcp, y=name, fill=name)) +
 geom_density_ridges(scale=.85)
```

```{r}
weather_df |> 
  ggplot(aes(x=name, y=prcp)) + 
  geom_boxplot()
```

```{r}
weather_df |> 
  filter(prcp >10, prcp < 1000) |> 
  ggplot(aes(x=prcp, fill=name)) +
  geom_density(alpha=.3)
```

## Saving and embedding plots

saving plots
```{r}
ggp_weather=
weather_df |> 
  ggplot(aes(x= date, y=tmax, color=name)) +
  geom_point()

ggsave("plots/ggp_weather.pdf", ggp_weather, width=8, height=6)
```

embedding plots

```{r, fig.width=6, fig.asp=.6}
weather_df |> 
  ggplot(aes(x= date, y=tmax, color=name)) +
  geom_point()
```

Vis II notes start..
Make a scatterplot but fancy this time 

```{r}
weather_df |> 
  ggplot(aes(x=tmin, y=tmax, color=name)) +
  geom_point(alpha=.3) +
  labs(
    title= "Temperature Scatterplot",
    x= "Minimum Temp (C)",
    y = "Maximum Temp (C)",
    color= "Location",
    caption= "Weather data taken ffrom rnoaa package for three stations"
  )
```

Scales--start with 'x' and 'y' and then do 'color'

```{r}
weather_df |> 
  ggplot(aes(x=tmin, y=tmax, color=name)) +
  geom_point(alpha=.3) +
  labs(
    title= "Temperature Scatterplot",
    x= "Minimum Temp (C)",
    y = "Maximum Temp (C)",
    color= "Location",
    caption= "Weather data taken ffrom rnoaa package for three stations"
  ) +
  scale_x_continuous(
    breaks = c(-15, 0, 20),
    labels = c("-15C", "0", "20")
  ) + 
  scale_y_continuous(
    limits = c(0, 30),
    transform = "sqrt"
  )
```

Look at color:

```{r}
weather_df |> 
  ggplot(aes(x=tmin, y=tmax, color=name)) +
  geom_point(alpha=.3) +
  labs(
    title= "Temperature Scatterplot",
    x= "Minimum Temp (C)",
    y = "Maximum Temp (C)",
    color= "Location",
    caption= "Weather data taken ffrom rnoaa package for three stations"
  ) +
  scale_color_hue(h= c(100,400)) + 
  viridis::scale_color_viridis(discrete = TRUE)
```

## Themes

```{r}
ggp_scatterplot=
  weather_df |> 
  ggplot(aes(x=tmin, y=tmax, color=name)) +
  geom_point(alpha=.3) +
  labs(
    title= "Temperature Scatterplot",
    x= "Minimum Temp (C)",
    y = "Maximum Temp (C)",
    color= "Location",
    caption= "Weather data taken ffrom rnoaa package for three stations"
  ) +
  scale_color_hue(h= c(100,400)) + 
  viridis::scale_color_viridis(discrete = TRUE)
```

```{r}
ggp_scatterplot + 
  theme(legend.position = "bottom")
```

Order matters..
making background black and white and then putting legend on bottom. if put legend position first wouldnt be on bottom bc themebw resets it. theme_minimal is like theme bw but without outside box
```{r}
ggp_scatterplot +
  theme_bw() +
  theme(legend.position = "bottom")
```

```{r}
ggp_scatterplot +
 theme(legend.position = "bottom") +
  ggthemes::theme_excel()
```

Learning assessment
```{r}
weather_df |> 
  ggplot(aes(x=date, y=tmax, color=name)) +
  geom_point(alpha=.3) +
  geom_smooth (se= FALSE) +
  viridis::scale_color_viridis(discrete = TRUE) +
  labs(
    title= "Temperature Plot",
    x= "Date",
    y= "Maximum Temperature (C)",
    size= "Precipitation"
  ) +
    theme_minimal() +
    theme(legend.position = "bottom")

```

Extra bonus stuff in 'ggplot'
Use diff datasets in diff 'geom's

```{r}
central_park_df=
  weather_df |> 
  filter(name== "CentralPark_NY")

molokai_df=
  weather_df |> 
  filter(name== "Molokai_HI")

molokai_df |> 
  ggplot(aes(x=date, y=tmax, color=name)) +
  geom_point() +
  geom_line(data= central_park_df)
```

## Multiple panels

```{r}
weather_df |> 
  ggplot(aes(x= tmax, fill=name)) +
  geom_density() +
  facet_grid(. ~ name)
```

```{r}
ggplot_tmax_tmin=
  weather_df |> 
  ggplot(aes(x=tmin, y=tmax, color=name)) +
  geom_point(alpha=.3)

ggp_tmax_density=
  weather_df |> 
  ggplot(aes(x= tmax, fill=name)) +
  geom_density(alpha=.3) 

ggp_tmax_date=
  weather_df |> 
  ggplot(aes(x=date, y=tmax, color=name)) +
  geom_point() +
  geom_smooth(se=FALSE)


(ggp_tmax_density + ggplot_tmax_tmin)/ ggp_tmax_date
```



## Data manipulation

```{r}

```



