---
title: "Vis I"
output: github_document
date: "2024-09-26"
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggridges)
```

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USW00022534", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2021-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = case_match(
      id, 
      "USW00094728" ~ "CentralPark_NY", 
      "USW00022534" ~ "Molokai_HI",
      "USS0023B17S" ~ "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())
```

Making out first plot :)

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) +
geom_point()
```
point says put points for all variables--making a scatterplot for us. aes is our aesthetics

```{r}
weather_df |> 
  ggplot(aes (x =tmin, y=tmax)) + 
  geom_point()
```
can also do it this way

```{r}
ggp_weather_scatterplot=
  weather_df |> 
   ggplot(aes (x =tmin, y=tmax)) + 
  geom_point()

ggp_weather_scatterplot
```

check why some rows are missing...
```{r}
weather_df |> 
  filter(is.na(tmax))
```
telling us there are 17 rows missing tmax


## Fancier scatterplots!

```{r}
weather_df |> 
  ggplot(aes(x=tmin, y=tmax, color=name)) +
  geom_point(alpha=.3, size=.8) + 
  geom_smooth(se = FALSE)
```
alpha makes points more transparent

Where you define aesthetics can matter--now putting color=name below and wont be applied to ggplot

```{r}
weather_df |> 
  ggplot(aes(x=tmin, y=tmax)) +
  geom_point(aes(color=name), alpha=.3, size=.8) + 
  geom_smooth(se = FALSE)
```

use faceting real quick

```{r}
weather_df |> 
  ggplot(aes(x=tmin, y=tmax, color=name)) +
  geom_point(alpha=.3) +
  geom_smooth(se= FALSE) + 
  facet_grid(. ~ name)
```
for faced grid . (nothing) goes in rows and name in columns. can reverse it too

Lets make a somewhat more interesting scatterplot 

```{r}
weather_df |> 
  ggplot(aes(x=date, y=tmax, color=name, size = prcp)) +
  geom_point(alpha=.3) +
  geom_smooth(se= FALSE) +
  facet_grid(. ~ name)
```

Learning assessment--need to manipulate data a little bit first

```{r}
weather_df |> 
  filter(name == "CentralPark_NY") |> 
  mutate(
    tmax_fahr = tmax + (9/5) + 32,
    tmin_fahr = tmin + (9/5) + 32) |> 
  ggplot(aes(x=tmin_fahr, y=tmax_fahr)) +
  geom_point() +
  geom_smooth(method= "lm", se=FALSE)
```




